### 2020-07-04 Thesis-Chapter2-Background.tex
Make more writing and editing improvements

\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack in which the suspect was killed, leaving behind a
locked iPhone, resulted in the first major battle of \ac{the-cw2}. The \ac{FBI} issued an order under the \acrlong{AWA}
to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
objected on grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
\cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the political world's attention from
February 16, 2016, when the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the
phone through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto
wars--terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
% FIXME: determine correct usage of -- vs ---
jointly released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not
make it to the floor, and the unresolved nature of the court case left the legal debate unresolved.

@@

The following sections list laws which directly or indirectly affect use encryption. \Ac{EA} is an international
problem, as digital technology has little respect for borders. This thesis is U.S.-focused, but briefly touches on the
regulatory environment of other countries.

@@

\subsubsection{Enacted U.S. Regulations}

@@

The most frequently cited constitutional amendments in the \ac{EA} debate. I: freedom of speech; IV: security against
unreasonable searches and seizures and requirements for warrants; V: right against self-incrimination.

@@

Amended the Wiretap Act, Stored Communications Act, and Pen Register Act. Provided explicit protections for private
electronic communications data, and updated definitions to contemporary technology. It granted law enforcement explicit
access to certain data as well; some terminology has aged poorly with technological evolution, making the access
increasingly permissive \cite{shamsi_2011}.

@@

The following are proposed laws and regulations that did not go into effect or have yet to be voted on:

@@

As described in \mysec{sec-history-cw1}, the \ac{clipperchip} was a voluntary \ac{EA} initiative from the White House.
The initiative targeted \ac{DIM} and established the Escrowed Encryption Standard, which includes an encrypted copy of
the session key in the message \cite{blaze_protocol_1994}. The initiative received a mixed reception and prompted much
debate \cite{kehl_right_2015}. When a security researcher discovered flaws that allowed users to subvert the
\ac{clipperchip} mechanisms, the proposal died \cite{blaze_protocol_1994}.

@@

The \ac{SAFEAct} would have broadly protected ``use any encryption regardless of the encryption algorithm selected,
encryption key length chosen, or implementation technique or medium used,'' excepting ``the unlawful use of encryption
in furtherance of a criminal act.'' It also would have lifted export controls. The bill grew support as it was
reintroduced in consecutive congresses, but was abandoned when the Clinton administration adopted the act's policies in
1999 \cite{kehl_right_2015}.

@@

The Senate's \ac{PROCODEAct} was very similar to House's \ac{SAFEAct}, though it focused primarily on removing export
controls, and used slightly weaker language protecting all forms of encryption. Like the \ac{SAFEAct}, it was abandoned
when the Clinton administration adopted the act's policies in 1996 \cite{thompson_2015}.

@@

\Ac{CESA} was a legislative proposal by Clinton administration introduced on same day as encryption policy changes that
led to the abandonment of the \ac{SAFEAct}. Section 102 lays out the government position precisely--encryption is an
important tool for information security, but is also used to ``hide unlawful activity by terrorists, drug traffickers,
child pornographers, and other criminals'' (the \ac{horsemen}), and therefore ``[a]ppropriate means must be available to
fulfill these law enforcement objectives.'' The bill would establish voluntary-participation third-party key escrow with
``recovery agents'' that would provide law enforcement with access to \ac{plaintext}. Access would be compelled through
a variety of mechanisms, including warrantless mechanisms through \ac{FISA}. The legislation never got a legislative
sponsor to be officially considered.

@@

Though this bill was not officially proposed, it represents one of the first legislative attempts of \ac{the-cw2} to
regulate encryption. This draft was released shortly after the close of the Apple vs. FBI case, and would have compelled
Apple to unlock the phone. Though it would not mandate a technical mechanism, the bill would have mandated that
manufacturers and service providers be able to provide \ac{EA}. This mandate notably goes beyond several regulatory
proposal in \ac{the-cw1}. Previewing a common phrase in recent \ac{EA} arguments \cite{geller_2019}, the bill's authors
both used the phrase ``above the law'' in promotion of the bill.

@@

Introduced by the chair of the Senate Judiciary Committee and one of the outspoken encryption critics during the 2019
Senate hearing on encryption and \ac{EA}, the \ac{EARNITAct}, designed to address \ac{CSAM}, has received attention as a
threat to encryption \cite{newman_2020} \cite{pfefferkorn_2020}. As described in \mysec{sec-history-current}, the bill
would revoke online platforms' liability protection for user-submitted content if is does not doing enough to combat
\ac{CSAM}. Critics fear that meeting the regulatory requirements would preclude \ac{E2EE}, and at least one popular
\ac{E2EE} messaging platform has threatened to pull out of the U.S. market if the bill were to pass \cite{lund_2020}.

@@

@@

France, Belgium, Germany, and Sweden require cooperation between telecommunications providers and law enforcement; for
France, Belgium, and Germany, that includes decrypting network traffic when possible. For Belgium, that decryption
capability is mandatory as of 2011. France, Germany, and Sweden each have some level of \ac{lawful-hacking} capability.
There is no E.U. legislation that requires key disclosure or decryption of network traffic.

@@

There are several technical approaches to \ac{EA}. This section breaks down these approaches according to their basic
approach to \ac{key-escrow}, to cryptographic recovery without a key, or to side-stepping the issue by using
non-technical means to acquire plaintext.

@@

The most obvious approach to \ac{EA} is to store additional copies of the encrypting keys, or information that can be
used to derive additional copies. This approach is called \ac{key-escrow}. There are several classes of key escrow.

@@

Some schemes compensate for the risk of concentrated sensitive information by massively distributing the secret
information. In this arrangement, trust is laid in the distrubuted system, not in any single party. Such systems offer
high data availability while making it difficult for access requests to be secret or to subvert access policies. This
approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has come into its own through the application of
blockchain \cite{phan_key_2017}. These approaches are not as developed as trusted-party \ac{key-escrow}.

@@

A third type of \ac{key-escrow} is device-based. In device-based \ac{key-escrow}, special key information is held not by
trusted parties or distributed systems, but on the device itself in specialized hardware. Advances in secure hardware
``enclaves'' since \ac{the-cw1} have enabled secure local key storage. Approaches relying on hardware primarily target
\ac{DAR}, and include time elements and split keys with trusted parties as well \cite{savage_lawful_2018}
\cite{ozzie_2018}.

@@

As introduced in \mysec{sec-reg-environment}, one workaround to direct \ac{EA} is \ac{lawful-hacking}. Instead of having
access to plaintext data via an alternative means of decryption, law enforcement is allowed to perform otherwise-illegal
hacking activity in order to find the key, compromise a device, or intercept plaintext versions of the data
\cite{kerr_encryption_2017}. While some see this as a viable middle ground \cite{nguyen_lawful_2017}, encouraging law
enforcement to exploit vulnerabilities creates mis-aligned incentives. If law enforcement opposes strong security in
general, it could foster the exploit market while costing ordinary users money and security \cite{soesanto_2018}.

@@

In some cases, a suspect in custody may have the ability to access the desired encrypted information via key, \ac{PIN},
or password. In these cases, authorities sometimes compel the suspect to disclose the password. In the U.S., this
approach draws debates about the fifth amendment right against self-incrimination \cite{bittenbender_2019}. Of course,
this approach cannot address \ac{DIM} or situations in which the knowledgeable party is not in custody.

@@

In order to advance the \ac{EA} debate, it helps to organize all the arguments used in a cohesive manner. One way to
organize arguments is to create argument maps. Argument maps are graphical representations of the logical structures and
relationships between statements, premises, and conclusions. Argument maps are an outgrowth of the same research by
Horst Rittel that originated the idea of \acp{wicked-prob}. In 1970 his research group developed \acp{IBIS} to break
down \acp{wicked-prob} and document the reasoned approaches that go into solving them \cite{kunz_issues_1970}. Interest
in \ac{IBIS} has waxed and waned. Some software tools have been developed, including the now-defunct gIBIS
\cite{conklin_gibis_1988} and Compendium \cite{dutoit_hypermedia_2006} tools designed to produce graphical
representations. Argdown is a recent tool that generates argument maps from structures specified in a Markdown-based
language \cite{voigt_argdown_2018}.

@@

\Acp{DFD} are one tool designed to address these questions. \Acp{DFD} feature processes, data flows, data stores, and
external entities. \Acp{DFD} well suited for threat modeling because security vulnerabilities tend to follow data flow,
not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is the
primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In this thesis I create \acp{DFD} in this
style to analyze the threats already present and those that \ac{EA} would introduce.
% TODO: reference forward chapter

>>>>

\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack in which the suspect was killed, leaving behind a
locked iPhone, resulted in the first major battle of \ac{the-cw2}. The \ac{FBI} issued an order under the \acrlong{AWA}
to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
objected on grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
\cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the political world's attention from
February 16, 2016, when the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the
phone through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto
wars--terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
% FIXME: determine correct usage of -- vs ---
% (https://www.chicagomanualofstyle.org/qanda/data/faq/topics/HyphensEnDashesEmDashes/faq0002.html)
jointly released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not
make it to the floor, and the unresolved nature of the court case left the legal debate unresolved.

@@

The following sections list laws which directly or indirectly affect use of encryption. Though this thesis is
U.S.-focused, it is worth briefly touching on the regulatory environment of other countries. \Ac{EA} is an international
problem, as digital technology has little respect for political borders.

@@

\subsubsection{Enacted U.S. Regulations}
% FIXME: distinguish statutes vs. regulations; rename "failed" section

@@

The most frequently cited constitutional amendments in the \ac{EA} debate. I: freedom of speech; IV: protection from
unreasonable searches and seizures and requirements for warrants; V: right against self-incrimination.

@@

Amended the Wiretap Act, Stored Communications Act, and Pen Register Act. Provided explicit protections for private
% FIXME: Briefly introduce these. Even just "Amended the switch-board-era"...
electronic communications data, and updated definitions to contemporary technology. It granted law enforcement explicit
access to certain data as well; some terminology has aged poorly with technological evolution, making the access
increasingly permissive \cite{shamsi_2011}.

@@

The following are proposed laws and regulations that did not go into effect or are still in process:

@@

As described in \mysec{sec-history-cw1}, the \ac{clipperchip} was a voluntary \ac{EA} initiative from the White House.
The initiative targeted \ac{DIM} and established the Escrowed Encryption Standard, which includes an encrypted copy of
the session key in the message \cite{blaze_protocol_1994}. The initiative prompted debate but received a largely hostile
reception \cite{kehl_right_2015}. When a security researcher discovered flaws that allowed users to subvert the
\ac{clipperchip} mechanisms, the proposal died \cite{blaze_protocol_1994}.

@@

The \ac{SAFEAct} would have broadly protected ``use any encryption regardless of the encryption algorithm selected,
encryption key length chosen, or implementation technique or medium used,'' excepting ``the unlawful use of encryption
in furtherance of a criminal act.'' It also would have lifted export controls. The bill grew support as it was
reintroduced in consecutive congresses, but was abandoned when the Clinton administration adopted the act's
pro-encryption policies in 1999 \cite{kehl_right_2015}.

@@

The Senate's \ac{PROCODEAct} was very similar to House's \ac{SAFEAct}, though it focused primarily on removing export
controls, and used slightly weaker language protecting all forms of encryption. Like the \ac{SAFEAct}, it was abandoned
when the Clinton administration adopted the act's export control policies in 1996 \cite{thompson_2015}.

@@

\Ac{CESA} was a legislative proposal by Clinton administration introduced on same day as encryption policy changes that
led to the abandonment of the \ac{SAFEAct}. Section 102 lays out the government position precisely--encryption is an
important tool for information security, but is also used to ``hide unlawful activity by terrorists, drug traffickers,
child pornographers, and other criminals'' (the \ac{horsemen}), and therefore ``[a]ppropriate means must be available to
fulfill these law enforcement objectives.'' The bill would establish voluntary-participation third-party key escrow with
``recovery agents'' that would provide law enforcement with access to \ac{plaintext}. Access would be compelled through
a variety of mechanisms, including warrantless mechanisms through \ac{FISA}. The White House failed to attract a
member of Congress to officially sponsor the bill.

@@

Though this bill was not officially proposed, it represents one of the first legislative attempts to regulate encryption
in \ac{the-cw2}. This draft was released shortly after the close of the Apple vs. FBI case, and would have compelled
Apple to unlock the phone. Though it would not mandate a technical mechanism, the bill would have mandated that
manufacturers and service providers be able to provide \ac{EA}. This mandate notably goes beyond several regulatory
proposals in \ac{the-cw1}.
% TODO: how so?
Previewing a common phrase in recent \ac{EA} arguments \cite{geller_2019}, the draft's
authors both used the phrase ``above the law'' in promotion of the bill.
% TODO: how does this differ from the new LEAD Act?

@@

Introduced by the chair of the Senate Judiciary Committee, one of the outspoken encryption critics during the 2019
Senate hearing on encryption and \ac{EA}, the \ac{EARNITAct} is designed to address \ac{CSAM} and has received attention
as a threat to encryption \cite{newman_2020} \cite{pfefferkorn_2020}. As described in \mysec{sec-history-current}, the
bill would revoke online platforms' liability protection for user-submitted content if it is does not doing enough to
combat \ac{CSAM}. Critics fear that meeting the regulatory requirements would preclude \ac{E2EE}, and at least one
popular \ac{E2EE} messaging platform has threatened to pull out of the U.S. market if the bill were to pass
\cite{lund_2020}.

@@

    \law{2020}{The \acf{LEADAct}}{graham_2020}{

Introduced by the same senator as the \ac{EARNITAct}, the \ac{LEADAct} is a direct attack on encryption. The
\ac{LEADAct} specifically addresses \ac{DAR} and \ac{DIM} by forcing device, operating system, and application
developers to implement the ability to decrypt all data stored or passing through the device or software system
``concurrently with their transmission'' ``unless the independent actions of an unaffiliated entity make it technically
impossible to do so.'' The act neither mandates nor recommends any technical approach to providing this capability.

% TODO: how does this differ from the old "Compliance with Court Orders Act" draft?

@@

France, Belgium, Germany, and Sweden require cooperation between telecommunications providers and law enforcement; for
France, Belgium, and Germany, that includes decrypting network traffic when possible. France, Germany, and Sweden each
have some level of \ac{lawful-hacking} capability. There is no E.U. legislation that requires key disclosure or
decryption of network traffic.

@@

There are several technical approaches to \ac{EA}. This section categorizes these approaches according to one of three
basic approaches---\ac{key-escrow}, cryptographic recovery without a key, or non-cryptographic means of acquiring
plaintext.

@@

The most obvious approach to \ac{EA} is to store additional copies of the encrypting keys. A variation is to store
information that can be used to derive additional copies. This approach is called \ac{key-escrow}. There are several
classes of key escrow.

@@

Some schemes compensate for the risk of concentrated sensitive information by massively distributing the secret
information. In this arrangement, trust is put in the distrubuted system, not in any single party. Such systems offer
high data availability while making it difficult for attempts to acquire keys to be secret or to subvert access
policies. This approach was introduced in \ac{the-cw1} \cite{goos_oblivious_1996} and has come into its own through the
application of blockchain \cite{phan_key_2017}. These approaches are not as thoroughly explored as trusted-party
\ac{key-escrow}.

@@

A third type of \ac{key-escrow} is device-based. In device-based \ac{key-escrow}, special key information is held not by
trusted parties or distributed systems, but on the device itself in specialized hardware. Advances in secure hardware
``enclaves'' since \ac{the-cw1} have enabled secure local key storage. Approaches relying on hardware primarily target
\ac{DAR}, and may include time elements and split keys with trusted parties as well \cite{savage_lawful_2018}
\cite{ozzie_2018}.

@@

As introduced in \mysec{sec-reg-environment}, one workaround to direct \ac{EA} is \ac{lawful-hacking}. Under
\ac{lawful-hacking}, instead of having access to plaintext data via an alternative means of decryption, law enforcement
is allowed to perform otherwise-illegal hacking activity in order to find the key, compromise a device, or intercept
plaintext versions of the data \cite{kerr_encryption_2017}. While some see this as a viable middle ground
\cite{nguyen_lawful_2017}, encouraging law enforcement to exploit vulnerabilities creates mis-aligned incentives. If law
enforcement opposes strong security in general, it could foster the exploit market while costing ordinary users money
and security \cite{soesanto_2018}.

@@

In some cases, a suspect in custody may have the ability to access the desired encrypted information via a key,
\ac{PIN}, or password. In these cases, authorities sometimes compel the suspect to disclose the password. In the U.S.,
this approach draws debates about the fifth amendment right against self-incrimination \cite{bittenbender_2019}. Of
course, this approach cannot address \ac{DIM} or situations in which the knowledgeable party is not in custody.

@@

The arguments used in the \ac{EA} debate are diverse and complex; in order to advance the \ac{EA} discussion, it would
help to organize all the arguments used in a cohesive manner. One way to organize arguments is to create argument maps.
Argument maps are graphical representations of the logical structures and relationships between statements, premises,
and conclusions. Argument maps are an outgrowth of the same research by Horst Rittel that originated the idea of
\acp{wicked-prob}. In 1970 his research group developed \acp{IBIS} to break down \acp{wicked-prob} and document the
reasoned approaches that go into solving them \cite{kunz_issues_1970}. Interest in \ac{IBIS} has waxed and waned. Some
software tools have been developed, including the now-defunct gIBIS \cite{conklin_gibis_1988} and Compendium
\cite{dutoit_hypermedia_2006} tools designed to produce graphical representations. Argdown is a recent tool that
generates argument maps from structures specified in a Markdown-based language \cite{voigt_argdown_2018}.

@@

\Acp{DFD} are one tool designed to address these questions. \Acp{DFD} feature processes, data flows, data stores, and
external entities. \Acp{DFD} are well suited for threat modeling because security vulnerabilities tend to follow data
flow, not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is
the primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In this thesis I create \acp{DFD} in
this style to analyze the threats already present and those that \ac{EA} would introduce.
% TODO: reference forward chapter
